---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---



```{r, echo=F, message=F, results='hide'}
library(dplyr)
library(ggplot2)
library(Hmisc)
Sys.setlocale("LC_ALL","English")
```



## Loading and preprocessing the data

```{r}
activity = read.csv(unzip('activity.zip'))
```



## What is mean total number of steps taken per day?

```{r}
stepsByDay = with(activity, tapply(steps, date, sum, na.rm=T))

qplot(stepsByDay, xlab='Total steps per day', binwidth=500)
```

* Mean total number of steps per day: `r mean(stepsByDay,na.rm=T)`

* Median total number of steps per day: `r median(stepsByDay,na.rm=T)`



## What is the average daily activity pattern?

```{r}
averageStepsByInterval = aggregate(steps ~ interval, activity, mean, na.rm=T)

ggplot(averageStepsByInterval, aes(x=interval, y=steps)) +
    geom_line() +
    xlab("Interval") +
    ylab("Average number of steps") 
```

Interval `r averageStepsByInterval$interval[averageStepsByInterval$steps==max(averageStepsByInterval$steps)]` contains the  maximum number of steps.

## Imputing missing values

Number of incomplete cases: `r sum(!complete.cases(activity))`

```{r}
activity.imputed = activity
activity.imputed$steps = impute(activity$steps,mean)

stepsByDay = with(activity.imputed, tapply(steps, date, sum, na.rm=T))

qplot(stepsByDay, xlab='Total steps per day (Imputed)', binwidth=500)
```

* Mean total number of steps per day: `r mean(stepsByDay,na.rm=T)`

* Median total number of steps per day: `r median(stepsByDay,na.rm=T)`



## Are there differences in activity patterns between weekdays and weekends?

```{r}
activity.imputed$dayType = ifelse(weekdays(as.Date(activity.imputed$date)) %in%
                                    c('Saturday','Sunday'),'weekend','weekday') 
  
averageStepsByInterval = aggregate(steps ~ interval + dayType, activity.imputed, mean, na.rm=T)

ggplot(averageStepsByInterval, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dayType ~ .) +
    xlab("Interval") +
    ylab("Average number of steps") 
```




